---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "READMEfigs/",
  cache = TRUE
)
```

# tsfeatures

[![Pending Pull-Requests](http://githubbadges.herokuapp.com/robjhyndman/tsfeatures/pulls.svg?style=flat)](https://github.com/robjhyndman/tsfeatures/pulls)

The R package *tsfeatures* provides methods for extracting various features from time series data.

## Installation

The **stable** version on R CRAN is coming soon.

You can install the **development** version from [Github](https://github.com/robjhyndman/tsfeatures) with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("robjhyndman/tsfeatures")
```

## Usage

### Hyndman, Wang and Laptev (ICDM 2015)

```{r yahoo, message=FALSE}
library(tsfeatures)
library(tidyverse)
library(anomalous)

# Compute the features used in Hyndman, Wang & Laptev (ICDM 2015).
# Note that crossing_points, peak and trough are defined differently 
# in the tsfeatures package than in the Hyndman et al (2015) paper. 
# Other features are the same. Using the real data from the paper
yahoo <- cbind(dat0, dat1, dat2, dat3)
hwl <- bind_cols(
         tsfeatures(yahoo,
           c("acf1","entropy","lumpiness",
             "flat_spots","crossing_points")),
         tsfeatures(yahoo,"stl_features", s.window='periodic', robust=TRUE),
         tsfeatures(yahoo, "max_kl_shift", width=48),
         tsfeatures(yahoo,
           c("mean","var"), scale=FALSE, na.rm=TRUE),
         tsfeatures(yahoo,
           c("max_level_shift","max_var_shift"), trim=TRUE)) %>%
  select(mean, var, acf1, trend, linearity, curvature, 
         seasonal_strength, peak, trough,
         entropy, lumpiness, spike, max_level_shift, max_var_shift, flat_spots,
         crossing_points, max_kl_shift, time_kl_shift)
```

```{r yahoo2, dependson="yahoo"}
# 2-d Feature space
prcomp(na.omit(hwl), scale=TRUE)$x %>% 
  as_tibble() %>%
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point()
```

### Kang, Hyndman & Smith-Miles (IJF 2017)

```{r ijf2017}
library(tsfeatures)
library(tidyverse)
library(forecast)


# Compute the features used in Kang, Hyndman & Smith-Miles (IJF 2017).
# Note that the trend and ACF1 are computed differently for non-seasonal
# data in the tsfeatures package than in the Kang et al (2017). 
# tsfeatures uses mstl() which uses supsmu() for the trend calculation with 
# non-seasonal data, whereas Kang et al used a penalized regression spline
# computed using mgcv instead.  Other features are the same.

M3data <- purrr::map(Mcomp::M3, 
  function(x){
      tspx <- tsp(x$x)
      ts(c(x$x,x$xx), start=tspx[1], frequency=tspx[3])
  })
khs_stl <- function(x,...)
{
  lambda <- BoxCox.lambda(x, lower=0, upper=1, method='loglik')
  y <- BoxCox(x, lambda)
  c(stl_features(y,s.window='periodic', robust=TRUE, ...), lambda=lambda)
}
khs <- bind_cols(
  tsfeatures(M3data, c("frequency", "entropy")),
  tsfeatures(M3data, "khs_stl", scale=FALSE)) %>% 
  select(frequency, entropy, trend, seasonal_strength, acfremainder, lambda) %>%
  replace_na(list(seasonal_strength=0)) %>%
  rename(
    Frequency = frequency,
    Entropy = entropy,
    Trend = trend,
    Season = seasonal_strength,
    ACF1 = acfremainder,
    Lambda = lambda) %>%
  mutate(Period = as.factor(Frequency))
```

```{r ijf2017graphs, message=FALSE, dependson="ijf2017"}
# Fig 1 of paper
khs %>% 
  select(Period, Entropy, Trend, Season, ACF1, Lambda) %>%
  GGally::ggpairs()

# 2-d Feature space (Top of Fig 2)
prcomp(select(khs, -Period), scale=TRUE)$x %>%
  as_tibble() %>%
  bind_cols(Period=khs$Period) %>%
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(col=Period))
```

### Kang, Hyndman & Li (in preparation)

```{r mars, message=FALSE}
# Compute all features in MARs paper

library(tsfeatures)
library(tidyverse)
library(forecast)

M3data <- purrr::map(Mcomp::M3, 
  function(x){
      tspx <- tsp(x$x)
      ts(c(x$x,x$xx), start=tspx[1], frequency=tspx[3])
  })

nsdiffs <- function(x){
  c(nsdiffs=ifelse(frequency(x)==1L, 1L, forecast::nsdiffs(x)))
}

features1 <- c(
  "entropy",
  "stl_features",
  "acf1",
  "max_kl_shift",
  "lumpiness",
  "ndiffs",
  "nonlinearity",
  "frequency",
  "nsdiffs",
  "heterogeneity"
)
features2 <- c(  
  "max_level_shift",
  "max_var_shift"
)

yk <- bind_cols(
  tsfeatures(M3data, features1),
  tsfeatures(M3data, features2, trim=TRUE)) %>%
  replace_na(list(
    seasonal_strength=0,
    max_kl_shift=0)) %>%
  map_dfr(rank)/length(M3data)

# Retain features that have meaning for M3
yk <- select(yk,
             seasonal_period, ndiffs, nsdiffs, entropy, Nonlinearity,
               trend, spike, seasonal_strength, linearity, curvature,
               acfremainder, acf1, max_kl_shift, lumpiness,
               ARCHtest.p, GARCHtest.p, Boxtest.p, GARCHBoxtest.p,
               Hetero, max_level_shift, max_var_shift)
```

```{r marsembed, message=FALSE, dependson="mars"}
# Create distance matrix
d <- dist(yk)
# Embed to 2d using tsne
ei <- embedding(d, method="Rtsne")
```
  
```{r marsgraphs, message=FALSE, dependson="marsembed"}
ei %>%
  bind_cols(tsfeatures(M3data, "frequency")) %>%
  mutate(Period = as.factor(frequency)) %>%
  ggplot(aes(x=Comp1, y=Comp2)) +
  geom_point(aes(col=Period))
```

```{r marsnnet, message=FALSE, warning=FALSE, dependson="marsembed"}
# Learn Neural net to model tsne embedding
library(caret)
df <- bind_cols(ei, yk)
nn1 <- avNNet(df[,-2],df$Comp1, repeats=10, size=5, linout=TRUE, trace=FALSE)
cor(predict(nn1),df$Comp1)

nn2 <- avNNet(df[,-1],df$Comp2, repeats=10, size=5, linout=TRUE, trace=FALSE)
cor(predict(nn2),df$Comp2)

# Now plot estimated components
tibble(Comp1=predict(nn1), Comp2=predict(nn2)) %>%
  bind_cols(tsfeatures(M3data, "frequency")) %>%
  mutate(Period = as.factor(frequency)) %>%
  ggplot(aes(x=Comp1, y=Comp2)) +
  geom_point(aes(col=Period))
```


## License

This package is free and open source software, licensed under GPL-3.
