---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "READMEfigs/",
  cache = TRUE
)
```

# tsfeatures

[![Pending Pull-Requests](http://githubbadges.herokuapp.com/robjhyndman/tsfeatures/pulls.svg?style=flat)](https://github.com/robjhyndman/tsfeatures/pulls)

The R package *tsfeatures* provides methods for extracting various features from time series data.

## Installation

The **stable** version on R CRAN is coming soon.

You can install the **development** version from [Github](https://github.com/robjhyndman/tsfeatures) with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("robjhyndman/tsfeatures")
```

## Usage (still in development)

### Hyndman, Wang and Laptev (ICDM 2015)

```{r yahoo, message=FALSE, eval=FALSE}
library(anomalous)

yahoo_real <- c(as.list(dat0),as.list(dat1),as.list(dat2),as.list(dat3))
yahoo_simulated <- c(as.list(dat4),as.list(dat5))

KLscore <- function(x){max_kl_shift(x, width=48)}
yahoo_features <- c("mean","var","acf1","stl_features","entropy",
                    "lumpiness","max_level_shift","max_var_shift","KLscore",
                    "flat_spots","crossing_points")
real_features <- tsfeatures(yahoo_real, yahoo_features) %>%
  select(mean, var, acf1, trend, linearity, curvature, season, peak, trough, entropy,
         lumpiness, spike, max_level_shift, max_var_shift, flat_spots, crossing_points,
         max_kl_shift, time_kl_shift)
sim_features <- tsfeatures(yahoo_simulated, yahoo_features) %>%
  select(mean, var, acf1, trend, linearity, curvature, entropy,
         lumpiness, spike, max_level_shift, max_var_shift, flat_spots, crossing_points,
         max_kl_shift, time_kl_shift)
```

### Kang, Hyndman & Smith-Miles (IJF 2017)

```{r ijf2017}
library(tsfeatures)
library(tidyverse)

M3data <- purrr::map(Mcomp::M3, function(x)x$x)
Lambda <- function(x){
  suppressWarnings(forecast::BoxCox.lambda(x, lower=0, upper=1))
}
khs <- tsfeatures(M3data, c("entropy","stl_features","frequency","Lambda")) %>% 
  select(frequency, entropy, trend, season, acfremainder, Lambda) %>%
  replace_na(list(season=1)) %>%
  rename(
    Period = frequency,
    Entropy = entropy,
    Trend = trend,
    Season = season,
    ACF1 = acfremainder) %>%
  mutate(Period = as.factor(Period))
```

```{r ijf2017graphs, message=FALSE}
# Fig 1 of paper
GGally::ggpairs(khs)

# 2-d Feature space (Top of Fig 2)
prcomp(khs[,-1])$x %>%
  as_tibble() %>%
  bind_cols(Period=khs$Period) %>%
  ggplot(aes(x=PC1, y=PC2)) +
    geom_point(aes(col=Period))
```

### Kang, Hyndman & Li (in preparation)

```{r mars, message=FALSE, eval=FALSE}
# Compute all features in MARs paper

library(tsfeatures)
library(tidyverse)
library(forecast)

M3data <- purrr::map(Mcomp::M3, function(x)x$x)

nsdiffs <- function(x){
  c(nsdiffs=ifelse(frequency(x)==1L, 1L, forecast::nsdiffs(x)))
}
features1 <- c(
  "entropy",
  "stl_features",
  "acf1",
  "max_kl_shift",
  "lumpiness",
  "ndiffs",
  "nonlinearity",
  "frequency",
  "nsdiffs",
  "heterogeneity"
)
features2 <- c(  
  "max_level_shift",
  "max_var_shift"
)

yk <- cbind(
        tsfeatures(M3data, features1),
        tsfeatures(M3data, features2, trim=TRUE)) %>% 
  select(entropy, trend, linearity, curvature, acf1,
    acfremainder, max_kl_shift, max_level_shift, max_var_shift, 
    ndiffs, Nonlinearity, frequency, season, nsdiffs,
    ARCHtest.p, GARCHtest.p, Boxtest.p, GARCHBoxtest.p, Hetero)
# What happened to var change on remainder?
```

## License

This package is free and open source software, licensed under GPL-3.
